---
title: "Abalone Exploration"
author: "Zachary Horton and Tanner Street"
date: "2023-12-12"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      fig.align = "center",
                      warning = FALSE, 
                      message = FALSE)
knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = 1.0
  )
knitr::opts_chunk$set(attr.source = '.numberLines')

#setwd("~/Documents/MIT/15.095_Machine_Learning/Project/ML-Project")
library(ggplot2)
library(caret)
library(readr)
library(readxl)
library(glmnet) # look at this package
library(dplyr)
library(tidyr)
library(lars) # look at this package
library(leaps) # look at this package
library(gbm)
library(corrplot)
library(lubridate)
library(car)
library(splines)
library(corrplot)
library(GGally)
library(caTools)
```

```{r, include=FALSE}
model.colors <- c(`K-Means` = "#005B41", 
                  `K-Medoids` = "#161A30",
                  `Opt. K-Means (Cold)` ="#427D9D", 
                  `Opt. K-Means (Warm)` = "#D80032", 
                  `Opt. K-Medoids (Cold)` = "#64CCC5", 
                  `Opt. K-Medoids (Warm)` = "#F05941")
text.colors <- c(`K-Means` = "white", 
                 `K-Medoids` = "white",
                 `Opt. K-Means (Cold)` ="white", 
                 `Opt. K-Means (Warm)` = "white", 
                 `Opt. K-Medoids (Cold)` = "black", 
                 `Opt. K-Medoids (Warm)` = "white")
update_geom_defaults("text", list(size = 3.5))

## Plotting Functions:
wcss_grouped <- function(data, transformation, proportion, distance, y_scale="fixed") {
    if (y_scale == "fixed") {
        g <- ggplot(data=data %>%
                   filter(Transformation == tolower(transformation),
                          Proportion == proportion,
                          `Optimizer Distance` == tolower(distance)), 
               aes(x=K, y=WCSS, group=Model, fill=Model)) +
            geom_col() +
            facet_grid(cols=vars(Model),
                       rows=vars(`Time Limit`)) +
            labs(title=paste0("WCSS: ", transformation, " Data | Proportion = ", proportion*100, "% | ", distance, " Optimizer")) +
            theme_classic() +
            theme(legend.position = "bottom") +
            scale_fill_manual(values=model.colors) +
            geom_text(aes(label = round(WCSS, digits=0), color=Model), show.legend=FALSE, vjust = 1.3) +
            scale_color_manual(values=text.colors)

        
        return(g)
    } else{
        g <- ggplot(data=data %>%
                   filter(Transformation == tolower(transformation),
                          Proportion == proportion,
                          `Optimizer Distance` == tolower(distance)), 
          aes(x=K, y=WCSS, group=Model, fill=Model)) +
        geom_col() +
        facet_grid(cols=vars(Model),
                   rows=vars(`Time Limit`),
                   scales="free_y") +
        labs(title=paste0("WCSS: ", transformation, " Data | Proportion = ", proportion*100, "% | ", distance, " Optimizer")) +
        theme_classic() +
        theme(legend.position = "bottom") +
        scale_fill_manual(values=model.colors) +
        scale_color_manual(values=model.colors) +
        geom_text(aes(label = round(WCSS, digits=0), color=Model), show.legend=FALSE, vjust = 1.3) +
        scale_color_manual(values=text.colors)        
        return(g)        
    }
}

wcss_individual <- function(data, transformation, proportion, distance, time, y_lower, y_upper) {
    g <- ggplot(data=data %>%
               filter(Transformation == tolower(transformation),
                      Proportion == proportion,
                      `Optimizer Distance` == tolower(distance),
                      `Time Limit` == time), 
        aes(x=K, y=WCSS, group=Model, fill=Model)) +
        geom_bar(stat="identity", position=position_dodge(1)) +
        labs(title=paste0("WCSS: ", transformation, " Data | Proportion = ", proportion*100, "% | ", time, " seconds | ", distance, " Optimizer")) +
        coord_cartesian(ylim = c(y_lower, y_upper)) +
        theme_classic() +
        theme(legend.position = "bottom") +
        scale_fill_manual(values=model.colors) +
        scale_color_manual(values=model.colors) +
        geom_text(aes(label = round(WCSS, digits=0), color=Model), show.legend=FALSE, vjust = 1.3, , position=position_dodge(1)) +
        scale_color_manual(values=text.colors)
    return(g)
}

silhouette_grouped <- function(data, transformation, proportion, distance, y_scale="fixed") {
    if (y_scale == "fixed") {
        g <- ggplot(data=data %>%
               filter(Transformation == tolower(transformation),
                      Proportion == proportion,
                      `Optimizer Distance` == tolower(distance)), 
        aes(x=K, y=Silhouette, group=Model, fill=Model)) +
        geom_col() +
        facet_grid(cols=vars(Model),
                   rows=vars(`Time Limit`)) +
        labs(title=paste0("Silhouette: ", transformation, " Data | Proportion = ", proportion*100, "% | ", distance, " Optimizer")) +
        theme_classic() +
        theme(legend.position = "bottom") +
        scale_fill_manual(values=model.colors)  +
        scale_color_manual(values=model.colors) +
        geom_text(aes(label = round(Silhouette, digits=0), color=Model), show.legend=FALSE, vjust = 1.3) +
            scale_color_manual(values=text.colors)    
        return(g)
    } else{
         g <- ggplot(data=data %>%
               filter(Transformation == tolower(transformation),
                      Proportion == proportion,
                      `Optimizer Distance` == tolower(distance)), 
           aes(x=K, y=Silhouette, group=Model, fill=Model)) +
        geom_col() +
        facet_grid(cols=vars(Model),
                   rows=vars(`Time Limit`),
                   scales="free_y") +
        labs(title=paste0("Silhouette: ", transformation, " Data | Proportion = ", proportion*100, "% | ", distance, " Optimizer")) +
        theme_classic() +
        theme(legend.position = "bottom") +
        scale_fill_manual(values=model.colors) +
        scale_color_manual(values=model.colors) +
        geom_text(aes(label = round(Silhouette, digits=2), color=Model), show.legend=FALSE, vjust = 1.3) +
        scale_color_manual(values=text.colors)    
    return(g)
    }
}

silhouette_individual <- function(data, transformation, proportion, distance, time, y_lower, y_upper) {
    g <- ggplot(data=data %>%
               filter(Transformation == tolower(transformation),
                      Proportion == proportion,
                      `Optimizer Distance` == tolower(distance),
                      `Time Limit` == time), 
        aes(x=K, y=Silhouette, group=Model, fill=Model)) +
        geom_bar(stat="identity", position=position_dodge(1)) +
        labs(title=paste0("Silhouette: ", transformation, " Data | Proportion = ", proportion*100, "% | ", time, " seconds | ", distance, " Optimizer")) +
        coord_cartesian(ylim = c(y_lower, y_upper)) +
        theme_classic() +
        theme(legend.position = "bottom") +
        scale_fill_manual(values=model.colors) +
        scale_color_manual(values=model.colors) +
        geom_text(aes(label = round(Silhouette, digits=2), color=Model), show.legend=FALSE, vjust = 1.3, position=position_dodge(1)) +
        scale_color_manual(values=text.colors)
    return(g)
}
```


## Read in Abalone Data Results 

```{r}
abalone_combined <- read_excel("./model_results/abalone_combined.xlsx")
abalone_combined <- abalone_combined %>%
                        filter(!WCSS ==  -15095.00)
abalone_combined$Model <- as.factor(abalone_combined$Model)
abalone_combined$`Time Limit` <- as.factor(abalone_combined$`Time Limit`)
abalone_combined$Proportion <- as.factor(abalone_combined$Proportion)
abalone_combined$Transformation  <- as.factor(abalone_combined$Transformation)
abalone_combined$K <- as.factor(abalone_combined$K)
abalone_combined$`Optimizer Distance` <- as.factor(abalone_combined$`Optimizer Distance`)
```

## WCSS Normalized Plots:

### Euclidean:

```{r}
wcss_grouped(abalone_combined, "Normalized", 0.10, "Euclidean", y_scale="free")
wcss_grouped(abalone_combined, "Normalized", 0.25, "Euclidean", y_scale="free")
wcss_grouped(abalone_combined, "Normalized", 0.75, "Euclidean", y_scale="free")
```

### Manhattan:

```{r}
wcss_grouped(abalone_combined, "Normalized", 0.10, "Manhattan", y_scale="free")
wcss_grouped(abalone_combined, "Normalized", 0.25, "Manhattan", y_scale="free")
wcss_grouped(abalone_combined, "Normalized", 0.75, "Manhattan", y_scale="free")
```

### Individual Euclidean:

```{r}
wcss_individual(abalone_combined, "Normalized", 0.10, "Euclidean", 30, 550, 2000)
wcss_individual(abalone_combined, "Normalized", 0.25, "Euclidean", 30, 1500, 7750)
wcss_individual(abalone_combined, "Normalized", 0.75, "Euclidean", 30, 5500, 10000)
wcss_individual(abalone_combined, "Normalized", 0.10, "Euclidean", 90, 550, 1750)
wcss_individual(abalone_combined, "Normalized", 0.25, "Euclidean", 90, 1250, 7500)
wcss_individual(abalone_combined, "Normalized", 0.75, "Euclidean", 90, 5000, 35000)
wcss_individual(abalone_combined, "Normalized", 0.10, "Euclidean", 180, 550, 1600)
wcss_individual(abalone_combined, "Normalized", 0.25, "Euclidean", 180, 1250, 6000)
wcss_individual(abalone_combined, "Normalized", 0.75, "Euclidean", 180, 5000, 45000)
```

### Individual Manhattan:

```{r}
wcss_individual(abalone_combined, "Normalized", 0.10, "Manhattan", 30, 4500, 19000)
wcss_individual(abalone_combined, "Normalized", 0.25, "Manhattan", 30, 10000, 76500)
wcss_individual(abalone_combined, "Normalized", 0.75, "Manhattan", 30, 35000, 220000)
wcss_individual(abalone_combined, "Normalized", 0.10, "Manhattan", 90, 4000, 8000)
wcss_individual(abalone_combined, "Normalized", 0.25, "Manhattan", 90, 10000, 70000)
wcss_individual(abalone_combined, "Normalized", 0.75, "Manhattan", 90, 35000, 210000)
wcss_individual(abalone_combined, "Normalized", 0.10, "Manhattan", 180, 4000, 8000)
wcss_individual(abalone_combined, "Normalized", 0.25, "Manhattan", 180, 10000, 55000)
wcss_individual(abalone_combined, "Normalized", 0.75, "Manhattan", 180, 35000, 210000)
```

## Silhouette Normalized Plots:

### Euclidean:

```{r}
silhouette_grouped(abalone_combined, "Normalized", 0.10, "Euclidean", y_scale="free")
silhouette_grouped(abalone_combined, "Normalized", 0.25, "Euclidean", y_scale="free")
silhouette_grouped(abalone_combined, "Normalized", 0.75, "Euclidean", y_scale="free")
```

### Manhattan:

```{r}
silhouette_grouped(abalone_combined, "Normalized", 0.10, "Manhattan", y_scale="free")
silhouette_grouped(abalone_combined, "Normalized", 0.25, "Manhattan", y_scale="free")
silhouette_grouped(abalone_combined, "Normalized", 0.75, "Manhattan", y_scale="free")
```

### Individual Euclidean:

```{r}
silhouette_individual(abalone_combined, "Normalized", 0.10, "Euclidean", 30, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.25, "Euclidean", 30, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.75, "Euclidean", 30, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.10, "Euclidean", 90, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.25, "Euclidean", 90, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.75, "Euclidean", 90, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.10, "Euclidean", 180, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.25, "Euclidean", 180, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.75, "Euclidean", 180, -0.05, 0.75)
```

### Individual Manhattan:

```{r}
silhouette_individual(abalone_combined, "Normalized", 0.10, "Manhattan", 30, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.25, "Manhattan", 30, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.75, "Manhattan", 30, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.10, "Manhattan", 90, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.25, "Manhattan", 90, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.75, "Manhattan", 90, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.10, "Manhattan", 180, -0.05, 0.75)
silhouette_individual(abalone_combined, "Normalized", 0.25, "Manhattan", 180, -0.05, 1.00)
silhouette_individual(abalone_combined, "Normalized", 0.75, "Manhattan", 180, -0.05, 0.70)
```

## WCSS Scaled Plots:

### Euclidean:

```{r}
wcss_grouped(abalone_combined, "Scaled", 0.10, "Euclidean", y_scale="free")
wcss_grouped(abalone_combined, "Scaled", 0.25, "Euclidean", y_scale="free")
wcss_grouped(abalone_combined, "Scaled", 0.75, "Euclidean", y_scale="free")
```

### Manhattan:

```{r}
wcss_grouped(abalone_combined, "Scaled", 0.10, "Manhattan", y_scale="free")
wcss_grouped(abalone_combined, "Scaled", 0.25, "Manhattan", y_scale="free")
wcss_grouped(abalone_combined, "Scaled", 0.75, "Manhattan", y_scale="free")
```

### Individual Euclidean:

```{r}
wcss_individual(abalone_combined, "Scaled", 0.10, "Euclidean", 30, 0, 200)
wcss_individual(abalone_combined, "Scaled", 0.25, "Euclidean", 30, 0, 750)
wcss_individual(abalone_combined, "Scaled", 0.75, "Euclidean", 30, 0, 500)
wcss_individual(abalone_combined, "Scaled", 0.10, "Euclidean", 90, 0, 150)
wcss_individual(abalone_combined, "Scaled", 0.25, "Euclidean", 90, 0, 1650)
wcss_individual(abalone_combined, "Scaled", 0.75, "Euclidean", 90, 0, 5750)
wcss_individual(abalone_combined, "Scaled", 0.10, "Euclidean", 180, 0, 100)
wcss_individual(abalone_combined, "Scaled", 0.25, "Euclidean", 180, 0, 1650)
wcss_individual(abalone_combined, "Scaled", 0.75, "Euclidean", 180, 0, 2300)
```

### Individual Manhattan:

```{r}
wcss_individual(abalone_combined, "Scaled", 0.10, "Manhattan", 30, 100, 320)
wcss_individual(abalone_combined, "Scaled", 0.25, "Manhattan", 30, 275, 3800)
wcss_individual(abalone_combined, "Scaled", 0.75, "Manhattan", 30, 0, 34000)
wcss_individual(abalone_combined, "Scaled", 0.10, "Manhattan", 90, 100, 250)
wcss_individual(abalone_combined, "Scaled", 0.25, "Manhattan", 90, 250, 1700)
wcss_individual(abalone_combined, "Scaled", 0.75, "Manhattan", 90, 1000, 16500)
wcss_individual(abalone_combined, "Scaled", 0.10, "Manhattan", 180, 100, 260)
wcss_individual(abalone_combined, "Scaled", 0.25, "Manhattan", 180, 250, 700)
wcss_individual(abalone_combined, "Scaled", 0.75, "Manhattan", 180, 950, 16200)
```

## Silhouette Scaled Plots:

### Euclidean:

```{r}
silhouette_grouped(abalone_combined, "Scaled", 0.10, "Euclidean", y_scale="free")
silhouette_grouped(abalone_combined, "Scaled", 0.25, "Euclidean", y_scale="free")
silhouette_grouped(abalone_combined, "Scaled", 0.75, "Euclidean", y_scale="free")
```

### Manhattan:

```{r}
silhouette_grouped(abalone_combined, "Scaled", 0.10, "Manhattan", y_scale="free")
silhouette_grouped(abalone_combined, "Scaled", 0.25, "Manhattan", y_scale="free")
silhouette_grouped(abalone_combined, "Scaled", 0.75, "Manhattan", y_scale="free")
```

### Individual Euclidean:

```{r}
silhouette_individual(abalone_combined, "Scaled", 0.10, "Euclidean", 30, -0.05, 0.75)
silhouette_individual(abalone_combined, "Scaled", 0.25, "Euclidean", 30, -0.10, 0.85)
silhouette_individual(abalone_combined, "Scaled", 0.75, "Euclidean", 30, -0.05, 0.75)
silhouette_individual(abalone_combined, "Scaled", 0.10, "Euclidean", 90, -0.05, 0.75)
silhouette_individual(abalone_combined, "Scaled", 0.25, "Euclidean", 90, -0.05, 0.80)
silhouette_individual(abalone_combined, "Scaled", 0.75, "Euclidean", 90, -0.05, 0.75)
silhouette_individual(abalone_combined, "Scaled", 0.10, "Euclidean", 180, -0.05, 0.75)
silhouette_individual(abalone_combined, "Scaled", 0.25, "Euclidean", 180, -0.05, 0.75)
silhouette_individual(abalone_combined, "Scaled", 0.75, "Euclidean", 180, -0.05, 0.75)
```

### Individual Manhattan:

```{r}
silhouette_individual(abalone_combined, "Scaled", 0.10, "Manhattan", 30, -0.05, 0.75)
silhouette_individual(abalone_combined, "Scaled", 0.25, "Manhattan", 30, -0.10, 0.85)
silhouette_individual(abalone_combined, "Scaled", 0.75, "Manhattan", 30, -0.05, 0.75)
silhouette_individual(abalone_combined, "Scaled", 0.10, "Manhattan", 90, -0.05, 0.75)
silhouette_individual(abalone_combined, "Scaled", 0.25, "Manhattan", 90, -0.05, 0.80)
silhouette_individual(abalone_combined, "Scaled", 0.75, "Manhattan", 90, -0.05, 0.95)
silhouette_individual(abalone_combined, "Scaled", 0.10, "Manhattan", 180, -0.05, 0.75)
silhouette_individual(abalone_combined, "Scaled", 0.25, "Manhattan", 180, -0.05, 0.75)
silhouette_individual(abalone_combined, "Scaled", 0.75, "Manhattan", 180, -0.10, 0.95)
```